Selective electroporation of the viral envelope via optimal control of transmembrane potential

### 2021-06-11 documents/lessons_learned.tex
nondimensionalizing isn't working
Hampath

@@

>>>>

it seems like older papers that establish the base facts of a field, perhaps because of un-searchable full text, or a lack of recent citations directly,means there is some sort of bias towards more recent, but more detail-sparse papers.

PLoS One doesn't allow footnotes. This seemed weird, but trying to rewrite inline actually does seem to make things a lot cleaner, and I'll try to keep this.

I've been muttering a lot about saving the results of the simulations. However, perhaps the real solution is to never undergo this trial-and-error technique - to learn enough about the problem, and to have enough testing in place. This might not ever be feasible.

@@

Memoization is more thoroughly considered in Scaling of Differential Equations with joblib.

### 2021-06-11 documents/pure_electroporation.tex
Hampath

	fortuitous Dielectric properties apparently due to the

@@

At the risk of spending more time producing underpowered "scientifically useless" results,

In a continued effort to make ultracrepidarian pseudo-physicists such as ourselves feel useful

@@

@@

It is, in general, something of a singular outlier in that it is one of very few biological mechanisms relevant at practical electric field intensities. Similar to thermal treatment, it also performs an operation something like integration with a threshold, a process distinct from normal absolute-value. It is perhaps this property which provides a small degree of tunability.

@@

The idea of inactivating viruses via electric fields is not new. Very recently, Osorio \cite{Receptor2021} have reported on a susceptibility of SARS-CoV-2's spike protein to intense electric fields; however, in general, techniques targeting proteins appear to require truly extreme; in their study, fields of up to 2.9 GV/m were required. Dissocation of capsid proteins has been predicted by Man \cite{Picosecond2016b} at 2 GV/m.

@@

There is a single report of the electroporation of HIV in the gray literature. This one of very few reports involving the electroporation of viruses, reporting the ~95\% inactivation of HIV using a standard commercial Bio-Rad electroporation system at 2.3 MV/m. However, the authors appear to have committed such an egregious breach of research ethics that we cannot cite them here.

@@

The most obvious difference is scale. Viruses are minute. All other properties of the system being equal, the standard Schwan equation for transmembrane voltage implies that the steady-state DC voltage is proportional to the size of the particle. In the case of a 100 nm virus

Madiyar's Vaccinia electroporation data aligns with. Assuming that the host becomes permeable at 0.05 MV/m, the 250 nm vaccinia might be expected to have a threshold of 4 MV/m. Mutatoyo's data also agree.

This does not agree with the two reports. It does not seem plausible that the

@@

Consider the following double gaussian waveform, each with a FWHM of 500 picoseconds, separated by 1 nanosecond. In this case, the peak virus transmembrane voltage is no longer 1/400 less than the host cell, but 1/6.5, and the total energy in each curve is also improved to a ratio of 1/37. The second, inverted pulse has the effect of quickly zeroing out the host cell's decaying potential to improve the virus' energy ratio.

@@

400x difference, reflected in the numerical steady state simulations.

@@

[step response]

And so, despite having similar time constants, a short gaussian pulse [] long

@@

The dielectric properties of viruses have been determined by the fascinating technique of electrorotation, dielectrophoresis, and dielectric spectroscopy. To prevent fluid heating, these are often conducted in low osmotic strength media. Enveloped viruses are permeable to small ions, and so some results may not necessarily representative of conductivities in physiological conditions; \cite{Assessment}.

@@

@@

We must apologize once again to the mathematicians for sins of terminology.

@@

@@

"Underactuated" robotics

@@

Kotnik discuss the transmembrane voltage due to several pulse shapes formed from the Laplace Transform of the original differential equation.

@@

@@

@@

\subsection*{}

@@

@@

@@

typeset in LaTeX using TeXstudio.

@@

>>>>

	Fortuitous dielectric properties ostensibly due to the densely-packed genome core and apparently only affecting some ssRNA viruses

@@

In a continued effort to make ultracrepidarian physicists such as ourselves feel useful in the wake of a purely biological threat, and at the risk of producing more underpowered and scientifically useless results,

@@

The idea of inactivating viruses via electric fields is not new. Very recently, Osorio \cite{Receptor2021} have reported on a susceptibility of SARS-CoV-2's spike protein to intense electric fields; however, in general, techniques targeting proteins appear to require truly extreme; fields of up to 2.9 GV/m were used in their computational study. Similarly, the dissocation of capsid proteins has been predicted by Man \cite{Picosecond2016b} at 2 GV/m.

@@

It is, in general, something of a singular outlier in that it is one of very few biological mechanisms that manifest at practical electric field intensities. Similar to thermal treatment, it also performs an operation something like integration with a threshold, a process distinct from normal absolute-value. It is perhaps this property which provides a small degree of tunability.

@@

@@

%There is a single report of the electroporation of HIV in the gray literature. This one of very few reports involving the electroporation of viruses, reporting the ~95\% inactivation of HIV using a standard commercial Bio-Rad electroporation system at 2.3 MV/m. However, the authors appear to have committed such an egregious breach of research ethics that we cannot cite them here.

@@

The most obvious difference is scale. All other properties of the system being equal, the standard Schwan equation for transmembrane voltage implies that the steady-state DC voltage is proportional to the size of the particle. In the case of a 100 nm virus, this would 400x difference, which is reflected precisely in the steady state numerical simulations that follow.

Madiyar's Vaccinia electroporation data aligns with. Assuming that the host becomes permeable at 0.05 MV/m, the 250 nm vaccinia might be expected to have a threshold of 4 MV/m. Mutatoyo's data also agree roughly.

This does not agree with Yamada \cite{Electroporation2012}'s data on electroporation of Hepatitis B.

@@

Consider the following bipolar gaussian waveform, each with a FWHM of 500 picoseconds, separated by 1 nanosecond. In this case, the peak virus transmembrane voltage is no longer 1/400 less than the host cell, but 1/6.5, and the total energy in each curve is also improved to a ratio of 1/37. The second, inverted pulse has the effect of quickly zeroing out the host cell's decaying potential to improve the virus' energy ratio.

@@

@@

@@

\section*{Empirical dielectric properties}

The dielectric properties of viruses have been determined by the fascinating and ingenious techniques involving electrorotation, dielectrophoresis, and dielectric spectroscopy. To prevent fluid heating, these are often conducted in low osmotic strength media; as enveloped viruses are permeable to small ions, some results may not necessarily representative of conductivities in physiological conditions or must be \cite{Assessment}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%55
\subsection*{Error in Kotnik}

\begin{tcolorbox}

	There appears to be a minor typographical error in equation 8, transfer function, from Kotnik \cite{Time1998}. The equation is given in terms of a and b coefficients,

	\textcolor{red}{$$F(s) = \frac{a_1 s^2 + a_2 s + a_3}{b_1 s^2 + b_2 s + b_3} \text{ (eq. 8)}$$}

	whereas the correct order appears to be:

	\textcolor{blue}{$$F(s) = \frac{a_3 s^2 + a_2 s + a_1}{b_3 s^2 + b_2 s + b_1}$$}

	The rest of the paper appears to obey this coefficient order: for instance, equation A6e in that paper, in the derivation of the step function:

	$$E(s) = \frac{1}{s}$$

	$$F(s) \cdot E(s) = \frac{a_3 s^2 + a_2 s + a_1}{b_3 s^2 + b_2 s + b_1} \frac{1}{s} = \frac{a_3 s^2 + a_2 s + a_1}{b_3 s^3 + b_2 s^2 + b_1 s} \text{ (eq. A6c)}$$ 	%confirmed with mathematica.

	We will continue to use this convention, assuming the subscripts on equations 9a through f are correct.

	This transposition was only detected late in the project, and so some documentation may still contain the incorrect coefficients. Beware.

\end{tcolorbox}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%55

@@

We must apologize once again to the mathematicians for sins of terminology and rigor.

@@

Scipy has a convenient LTI class that accepts the terms.

@@

underactuated robotics, simultaneous stabilization\cite{results1984}\cite{Simultaneous2005}

@@

Kotnik discuss the transmembrane voltage due to several pulse shapes constructed simply from the Laplace Transform of the original differential equation.

@@

Pole placement

\subsection*{Nonlinear model predictive control, quadratic regulators, }

\subsection*{Frequency-domain solution}

@@

$$ x_v(t) = A_v \frac{d^2u(t)}{dt^2} + B_v \frac{du(t)}{dt} + C_v u(t) - D_v \frac{dx_v(t)}{dt} - \frac{d^2 x_v(t)}{dt^2} $$

$$ x_h(t) = A_h \frac{d^2u(t)}{dt^2} + B_h \frac{du(t)}{dt} + C_h u(t) - D_h \frac{dx_h(t)}{dt} - \frac{d^2 x_h(t)}{dt^2} $$

@@

A system of first-order Differential-Algebraic Equations:

Smooth noise-free combined filter and differentiator.

An advantage of bang-bang or bang-off-bang over is that the values of u0 are not free. In this case, it was often found that (perhaps due to poorly-chosen values of the penalty factor rho, a poor initial guess, or some other issue in the problem setup), the gradient descent would decrease the control magnitude to a value below numerical accuracy. This necessitated a Lagrange constraint on the total, further increasing the computational complexity and restricting the class of solutions.

\subsection*{Analytic solution}

@@

\begin{tcolorbox}
	First instinct might be to try to solve for the waveform that sets alpha beta gamma to zero. 
\end{tcolorbox}

@@

"Bulk effects", as the concentration increases, can completely alter the. It may not be possible to produce a robust enough scheme

@@

typeset in LaTeX on TeXstudio and Gummi.

@@

The book Scaling of Differential Equations is an excellent guide to the numerical treatment of DEs.

### 2021-06-11 documents/scan_converter.tex
Oh hey the brilluoin just had a bug
whatever
Rewording text slightly in case biorxiv is a good idea
Pure electroporation
iLQR
Hampath

@@

@@

@@

@@

>>>>

\section{Scan converter}

@@

Gating time. Photoconductive gating is a particularly interesting technique for this.

@@

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=0.6\textwidth]{sensor_damage.jpg}
	\caption{}
\end{figure}

@@

Low vacuum, gas focused tubes are less predictable and more difficult to model. XOOPIC; the standard PIC code WARP has undocumented feature of collisional ionization but Tegerup has a guide to use. Simulation

PIC-LAS

@@

stray light

old canal ray paper

the old WARP has a module for collisional ionization, though it's got a bug \cite{Simulation}.
